\documentclass[8pt, letterpaper]{extarticle}

\usepackage[margin=0.1cm]{geometry}
\usepackage{amsmath}
\usepackage{graphicx}

\begin{document}

\noindent
\begin{minipage}[t]{0.45\textwidth}

\textbf{Graphs}

Assuming no self-loops, no multiple edges.

Undir: $m \leq \binom{n}{2},\ \sum_{v}deg(v) = 2m$

Dir: $m \leq n*(n-1),\ \sum_{v}indeg(v) = \sum_{v}outdeg(v) = m$

Subgraph: result of removing an edge, $V' \subseteq V, E' \subset E$

Induced SG: result of removing node. Is subgraph, $e \in E' \leftrightarrow (e \in E \wedge (u, v)) \in V$.

$u$ connected to $v$: $(u \sim v) \rightarrow \exists$ path from $u$ to $v$

$G$ connected if $(u \sim v)\ \forall (u, v)$


\textbf{Representations}

Adj. Matrix: row 1 = outgoing edge, col 1 = incoming edge

- Undir: $G = G^T$

- Find a neighbour: $O(n)$

- Access($v$) = $O(1)$, traverse $V = O(n)$

- Traverse all edges of $v = O(n)$

- Traverse all edges of $G = O(n^2)$ 

- $O(V^2) = O(n^2)$ space

Adj. List: $\sum_{v} outdeg(v) = E$

- $O(V + E) = O(n + m)$ space

- Find a neighbour: $O(1)$

- Traverse all edges of $v = O(|neighbours(v)|)$

- Traverse all edges of $G = O(m)$ 

- Access($v$) = $O(1)$, traverse $V = O(n)$


\textbf{Trees}

Tree: connected, acyclic graph

- Add edge $\rightarrow$ cycle; Remove edge $\rightarrow$ not connected

- n - 1 edges

Forest: graph with trees as connected components

Spanning tree: $T \subseteq G$ S.T $V(T) = V(G)$

Undir. is a tree $\leftrightarrow$ connected and E = V - 1


\textbf{BFS}

- $O(n + m)$ time (list), $O(n^2)$ (matrix)

- $\Theta(n)$ space for both list and matrix

- Finds all nodes, finds shortest path from $s$ to all others

\textbf{DFS}

- $O(n + m)$ time (list), $O(n^2)$ (matrix)

- $O(n + m)$ space

- Finds all nodes, finds shortest path from $s$ to all others

\textbf{Edge Classes}

The edge $(u, v)$ refers to the FOREST, not the graph.

- Tree: $(u, v) \in$ Forest

- Forward: $v$ is descendant of $u$

- Back: $v$ is ancestory of $u$ (back = fwd in undirs)

- Cross: none of the above are true.

- Und: DFS tree/fwd only, BFS tree/cross only

- Dir: DFS: all edges, BFS: no fwd edges


\textbf{DAGs}

- Source: no incoming edges

- Sink: no outgoing edges

- Multiple sources/sinks are possible. At least one of each in every DAG.

- Source \& Sink $\rightarrow$ no cycles

- A Digraph is a DAG $\leftrightarrow$ DFS has no back edges


\textbf{Toposort}

- Produces ordering s.t $(u,v) \in E \rightarrow u$ appears before $v$ in the ordering

- Digraph has a Toposort $\leftrightarrow$ it is a DAG

- Run DFS, order in decreasing order of finish time

- $O(n + m)$

- Orders are not unique.


\textbf{SCCs}

- Run DFS. Run DFS on $G^T$, in decreasing order of ftime. The forests of the transpose DFS are the SCCs.

- Independent of toposort ordering

\end{minipage}
\vline
\hspace{1pt}
\begin{minipage}[t]{0.45\textwidth}

    
\textbf{Minimum Spanning Trees}

- Spanning tree w/ minumum weight

- DFS and BFS build spanning trees, but \emph{not} necessarily the MST.

- Optimal Substructure: if $T = \text{MST of } G \rightarrow T[U], \text{ where } U \subset V$, if $T[U]$ is connected it is an MST.


\textbf{Prim: } add best vertex


\textbf{Kruskal's Algorithm} $(O((m + n)\log(n)))$

- Only consider edges that do NOT create a cycle (safe edges)

- Add lowest-cost edge on each iteration

- Has greedy-choice property


\textbf{Edge Switching}

- Let $e' \notin T$, where $T \cup \{e'\}$ has a cycle. For any $e \in E$, where $e$ is in the cycle, $T \cup \{e'\} - \{e\}$ is a tree.

\textbf{Single-Source Shortest Paths (SSSPs)}

- The problem: find the min-cost path from source $s$ to each vertex $v$.

- Shortest path is at most of length $n-1$.

\textbf{Relax}

\emph{If} $d(v) > d(u) + w(u, v)$, \emph{then} $d(v) \leftarrow d(u) + w(u, v)$

\textbf{Dijkstra}

It's kind of like $A^*$ search!

{\small \textbf{Algo:}} \emph{Init.} $S = \emptyset$; \emph{Init.} $d(v) = \infty,\ \forall v \neq s$, $d(s) = 0$; \emph{Init.} Min-Queue, keyed by $d(v)$. \emph{While} $Q \neq \emptyset$, \emph{extract} $u = \text{min}(Q)$, \emph{Add} $u$ to $S$; \emph{Relax} all edges $(u,v)$.

- List + binary heap (as a min-PQ): $O((m+n)\log(n))$

- Matrix: $O(n^2 + m\log(n))$

- Fib heap: $\Theta(m + n\log(n))$

- No negative cycles (undir), no negative edges (dir)

\textbf{Properties}

\emph{Subpath Optimality}: If $P = (s,...u,...v,...t)$, then each subpath is the shortest corresponding path.

\emph{Triangle Inequality}: $d(u,w) \leq d(u,v) + d(v,w)$.

\textbf{Relaxing an Edge}

- $d(v)$ only updated w/ relax $\rightarrow d(v) \geq d(s, v)$

relax($u, v$): if $d(v) > d(u) + w(u, v)$:

\hspace{0.5cm} $d(v) \leftarrow d(u) + w(u, v)$

\textbf{Bellman-Ford} $O(VE)$

- Can handle negative edge weights (but not negative cycles!)

- Updates distances OFFLINE.

{\small \textbf{Algo:}} \emph{Init.} $d(v) = \infty,\ \forall v \neq s,\ d(s) = 0$; \emph{For} $i \in \{1,...,n-1\}$, \emph{For} $(u,v) \in E$, \emph{Relax} edges $(u,v)$. \emph{Endfor.} \emph{For} $(u,v) \in E$; \emph{If } $d(u) + w(u,v) \leq d(v)$, \emph{Return} FALSE. \emph{Else} return TRUE.

- Use a DP table, dim = $0:(n-1) \times n$

- $d_{i}[v] \leftarrow \min(d_{i-1}[v], \min_{u \in Neigh(v)}[d_{i-1}[u] + w(u,v)])$

- Order irrelevant to final product.

\textbf{Floyd-Warshall} $O(n^3)$

- $V \times V$ table, $v_1,...v_n$.

- Allows negative weights, no negative cycles.

- All-pairs shortest path (APSP)

Suppose $V = \{1, 2, \ldots, n\}$. 

Then, $d[i,j,k] = $ distance of shortest path from $i$ to $j$, such that all intermediate vertices $\in \{1, 2, \ldots, k\}$.

- Case 1: don't need vertex $k$: $d[i,j,k] = d[i,j,(k-1)]$.

- Case 2: need vertex $k$: $d[i,j,k] = d[i,k,(k-1)] + d[k,j,(k-1)]$.

- So, $d[i,j,k] = \min(\text{Case 1, Case 2})$

- Base cases: $d[i,j,0] = w(i,j), \ d[i,i,k] = 0$.


\end{minipage}

\newpage

\noindent
\begin{minipage}[t]{0.45\textwidth}

    
\textbf{Greedy Algorithms}

Optimal Substructure: an optimal choice must be included in \emph{an} optimal solution, but not necessarily all of them.


\textbf{Fractional Knapsack}

$v_i = $ profit per unit weight $b_i / w_i$

$x_i = \min(w_i, W_\text{remaining})$

add $x_i$ amount of item $i$

$W_\text{remaining} \leftarrow W_\text{remaining} - x_i$

Choose item with highest $v_i$

$X^*$ is unique $\rightarrow$ all $x^* \in X^*$ are saturated.


\textbf{Job Scheduling}

- Earliest-Start-Time-First (ESTF)

Sort the $n$ jobs by start time $O(n\log(n))$; iterate $n$ times, and on each iteration find the right machine, schedule the job, and update the times. Total runtime: $O(n\log(n))$.

Alternatively use a min-heap: also $O(n\log(n))$.


\textbf{Activity Selection}

- Earliest-Finish-Time-First (EFTF); $\Theta(n\log(n))$

\textbf{Dynamic Programming}

\textbf{Integral Knapsack}

- $n$ items: table with $0:n-1$ rows, $0:W$ columns. $D$ is remaining capacity, $W$ is total capacity.

- $\forall D, \ A[0,D] = 0$

- $\forall i, \ A[i,0] = 0$

- Else, $A[i,D] = \max(A[i-1, D], v_i + A[i-1, D - w_i])$

\textbf{Rod Cutting}

- $r_0 = 0$

- Else, $r_n = \max_{i \in 1:n} (p_{i} + r_{n-i})$

\textbf{Longest Common Subsequence}

- $O(nm)$

- $len(X) = n, \ len(Y) = m$.

- Table: $0:n \times 0:m$.

- $A[i,j] = \max(D[i-1, j], D[i, j-1], 1 + D[i-1, j-1])$

- Last one only if $X[i] = Y[j]$.

\textbf{MCM}

- $n$ matrices

- $d \in \{0, \ldots, n+1\}$

- $A[i, j] = \min_{i \leq k < j} (A[i,k] + A[k+1, j], +\ d_{i-1}d_k d_j )$

\end{minipage}
\vline
\hspace{1pt}
\begin{minipage}[t]{0.45\textwidth}

\textbf{Asymptotic Cheatsheet}

For some set $S$,

If $f(n) \in S(g(n))$ and $g(n) \in S(h(n))$ then $f(n) \in S(h(n))$.

$\sum_{k=0}^{n} r^k = \frac{1 - r^{n+1}}{1 - r} = \frac{r^{n+1} - 1}{r - 1}$

$n! = \sqrt{2\pi n}(\frac{n}{e})^n$

{\small \textbf{Limits: }}

$\lim_{n \to \infty} \frac{h(n)}{f(n)} = 0 \rightarrow h(n) \in o(f(n))$

$\lim_{n \to \infty} \frac{h(n)}{f(n)} = k > 0 \rightarrow h(n) \in \Theta(f(n))$

$\log^k(n) \in o(n^{\varepsilon})$, $\forall k, \varepsilon > 0$

$n\log(n) \in \Theta(\log(n!))$

\textbf{Master Theorem}

Let $T(n) = aT(\frac{n}{b}) + f(n)$

Bottom-Heavy: Leaves dominate runtime

- $f(n) \in O(n^{\log_b(a) - \varepsilon}) \rightarrow T(n) \in \Theta(n^{\log_b(a)})$

Balanced: Leaves and internal nodes do equal work

- $f(n) \in \Theta(n^{\log_b(a) \log^k_n}) \rightarrow T(n) \in \Theta(n^{\log_b(a)}\log^{k+1}(n))$, $k \geq 0$

Top-Heavy: Eariler nodes dominate runtime

- $f(n) \in \Omega(n^{\log_b(a) + \varepsilon})$, and $af(\frac{n}{b}) \leq \delta f(n)$, $\delta < 1 \rightarrow T(n) \in \Theta(f(n))$

\textbf{Heaps}

Max-Heap Property: $A[\text{Parent}(i)] \geq A[i]$

Heap of $n$ keys has height $\lfloor \log(n) \rfloor$

buildMaxHeap: $O(n)$

extractMax: $O(\log(n))$

heapSort: $\Theta(n\log(n))$

\textbf{QuickSort}

- Partition: $O(n)$, returns index of pivot.

- Worst case: $O(n^2)$

- Average and best case: $O(n\log(n))$

- For any split of constant ratio, $\Theta(n\log(n))$

{\small \textbf{Randomized QuickSort}}

- Pivot selected uniformly at random

- Average, best, and expected worst-case runtime: $\Theta(n\log(n))$

\textbf{Sorting Lower Bound}

- Any comparison-based sorting algorithm requires $\Omega(n\log(n))$ comparisons in the worst case.

- A binary tree with $t$ leaves has at least $1 + \log(t)$, or eqiuvalently, a height of at least $\log(t)$.

- So in other words, a decision tree has at least $n!$ leaves. It follows that the height of the tree is at least $\log(n!)$.

- Height = edges

- levels = edges + 1

\textbf{Binary Search Trees}

- Right-(Left)-Rotate: $x$ becomes the right (left) child of new root, usually its left (right) child.

\textbf{AVL Trees}

- Height-balanced: $|h_l - h_r| \leq 1$

- $h = O(\log(n))$ 

\end{minipage}

\end{document}